Shapes analysis for time series.

### 2024-10-14 NEURIPS24/paper.tex
update
up

%Reproducing Kernel Hilbert Space (RKHS)
\vspace{-1ex}
\section{Background on LDDMM}
\vspace{-1ex}
\label{section:LDDMM}

In this part, we expose how to
 learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
 In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).

@@

With this in mind, the velocity field $v$ fitting the data can be
estimated by minimizing 
$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
 However, two computational challenges arise.
  First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
   In addition, a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, is sought to efficiently solve this minimization problem. 
\paragraph*{From deformations to geodesics.}
%\paragraph*{Regularizing the problem to derive uniqueness.}
It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$.
Following this geometric point of view, geodesics can be constructed on $\msh$ by using the following squared norm 
 \begin{equation}
  \label{eq:geodesics_original}
    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv\dd \tau
 \end{equation}
By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy-Lipschitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
Denoting by $\tau \mapsto \rho_{v_0}(\tau)\in\msh$ the geodesic starting from the identity with inital velocity $v_0\in \msv$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \rho_{v_0}(1)$. Using $\varphi^{\{v_0\}}$ instead of $\phi^v$, the previous matching problem becomes a \textit{geodesic shooting problem}:
 \begin{equation}
  \label{eq:geodesics_shooting}
  \inf_{v_0 \in \msv} \mathscr{L}(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}).
 \end{equation}
 Using $\varphi^{\{v_0\}}$ instead of $\phi^v$ for any $v\in \rml^2([0,1],\msv)$ regularizes the problem and induces a sparse representation for the learning diffeomorphisms.
 Moreover, by setting $\msv$ as an RKHS, the geodesic shooting problem has a unique solution and becomes tractable, as described in the next section.

@@

@@

\section{Related Works}
\vspace{-1ex}
Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
 The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
 Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
 The goal is to find a well-suited Riemannian structure according to the nature of the studied object.

@@

>>>>

%Reproducing Kernel Hilbert Space (RKHS)
\vspace{-1ex}
\section{Background on LDDMM}
\vspace{-1ex}
\label{section:LDDMM}
\sam{Shape-FPCA employs the Square Root Velocity Field (SRVF) representation to separate space and time. However, this method is designed for continuous objects and only applies to time series of the same length. This issue can be addressed through interpolation, but this approach is not always reliable in sparse and irregular sampling scenarios. Most FDA papers [2,3,4] that we are aware of cope with the same issue using interpolation or basis function expansion. In a nutshell, FDA methods have to deal with continuous objects, while LDDMM algorithms can keep a discrete-to-discrete analysis. }
In this part, we expose how to
 learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
 In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).

@@

With this in mind, the velocity field $v$ fitting the data can be
estimated by minimizing 
$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
 However, two computational challenges arise.
  First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
   In addition, a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, is sought to efficiently solve this minimization problem. 
\paragraph*{From deformations to geodesics.}
%\paragraph*{Regularizing the problem to derive uniqueness.}
It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$.
Following this geometric point of view, geodesics can be constructed on $\msh$ by using the following squared norm 
 \begin{equation}
  \label{eq:geodesics_original}
    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
 \end{equation}
By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy-Lipschitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
Denoting by $\tau \mapsto \rho_{v_0}(\tau)\in\msh$ the geodesic starting from the identity with inital velocity $v_0\in \msv$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \rho_{v_0}(1)$. Using $\varphi^{\{v_0\}}$ instead of $\phi^v$, the previous matching problem becomes a \textit{geodesic shooting problem}:
 \begin{equation}
  \label{eq:geodesics_shooting}
  \inf_{v_0 \in \msv} \mathscr{L}(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}).
 \end{equation}
 Using $\varphi^{\{v_0\}}$ instead of $\phi^v$ for any $v\in \rml^2([0,1],\msv)$ regularizes the problem and induces a sparse representation for the learning diffeomorphisms.
 Moreover, by setting $\msv$ as an RKHS, the geodesic shooting problem has a unique solution and becomes tractable, as described in the next section.

@@

A pedagogical \href{https://tslddmmapp.streamlit.app}{\underline{online application}} is available to inspect the effect of hyperprameters on geodesic shooting \eqref{eq:integration} and registration \eqref{eq:relaxation}.

@@

\section{Related Works}
\vspace{-1ex}

Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
 The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
 Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
 The goal is to find a well-suited Riemannian structure according to the nature of the studied object.

@@

 LDDMM framework has weaker assumptions than the Metamorphosis framework \cite{blanz2003face}. The 3DMM framework requires that each mesh be re-parametrized into a consistent form where the number of vertices, triangulation, and the anatomical meaning of each vertex are consistent across all meshes, as stated in the introduction of \cite{booth20163d}.
 In our context, we do not need such pre-processing; the time series graph can have different sizes.
